---
title: "Machine Learning Assignment 2"
output:
  pdf_document:
    toc: yes
  html_notebook:
    highlight: textmate
    theme: cerulean
  html_document:
    df_print: paged
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
dir.create("images")
```

***

# k-NN

Install packages if necessary. Uncomment before running.

```{r}
#install.packages("caret")
library(caret)

# install.packages("ISLR") # only install if needed
library(ISLR)

library(FNN)

#install.packages("dummies")
library(dummies)
library(gmodels)

```


Loading the data file and reviewing data structures

```{r}
UniversalBank<-read.csv("UniversalBank.csv")
head(UniversalBank)
summary(UniversalBank)
```

Excluding ID and Zip Code and selecting the data set with the mentioned criteria

```{r}
UniversalBank2<-UniversalBank[,c(-1,-5)]
UniversalBank1<-UniversalBank2[which(UniversalBank2$Age ==40 | UniversalBank2$Experience==10 | UniversalBank2$Income==84 | UniversalBank2$Family==2 | UniversalBank2$CCAvg==2 | UniversalBank2$Education==1 | UniversalBank2$Mortgage==0 | UniversalBank2$Securities.Account==0 | UniversalBank2$CD.Account==0 | UniversalBank2$Online==1 | UniversalBank2$CreditCard==1),]
str(UniversalBank1)
```


Creating Dummy Variables

```{r}
levels(UniversalBank1$Education)
dummy_model <- dummyVars(~Education,data=UniversalBank1)
head(predict(dummy_model,UniversalBank1))
```
Creating New Data Frame with the Data variables and doing the Data Normalization

```{r}
UBank<-dummy.data.frame(UniversalBank1, names = c("Education"), sep=".")
str(UBank)

norm_model<-preProcess(UBank, method = c('range'))
UBank_normalized<-predict(norm_model,UBank)
UBank_Predictors<-UBank_normalized[,-10]
UBank_labels<-UBank_normalized[,10]
```

Doing the data partition of 60% Training and 40% Validation

```{r}
set.seed(15)
inTrain = createDataPartition(UBank_normalized$Personal.Loan,p=0.6, list=FALSE)
Train_Data = UBank_normalized[inTrain,]
Val_Data = UBank_normalized[-inTrain,]
dim(Train_Data)
summary(Train_Data)
summary(Val_Data)

```




```{r}
Train_Predictors<-Train_Data[,-10]
Val_Predictors<-Val_Data[,-10]
Train_labels <-Train_Data[,10]
Val_labels <-Val_Data[,10]
Train_labels=as.factor(Train_labels)
Val_labels=as.factor(Val_labels)
UBank_labels<-as.factor(UBank_labels)

knn.pred <- knn(Train_Predictors,Val_Predictors,cl=Train_labels,k=1,prob = TRUE)
Q1 <- data.frame(40, 10, 84, 2, 2, 0, 1, 0, 0, 0, 0, 1, 1)
knn.pred1 <- knn(Train_Predictors, Q1, cl=Train_labels, k=1, prob = 0.5)
knn.pred1

```


```{r}

library(caret)
accuracy.df <- data.frame(k = seq(1, 14, 1), accuracy = rep(0, 14))
for(i in 1:14) {
knn <- knn(Train_Predictors, Val_Predictors, cl = Train_labels, k = i)
accuracy.df[i, 2] <- confusionMatrix(knn, Val_labels)$overall[1]
}
accuracy.df

```

```{r}
which.max( (accuracy.df$accuracy) )
```
Optimal value is of K=1 

#Confision Matrix
```{r}
knn.pred3 <- knn(Train_Predictors,Val_Predictors,cl=Train_labels,k=3,prob = TRUE)
confusionMatrix(knn.pred3,Val_labels)
```

Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above.

```{r}
set.seed(15)
Bank_Partition = createDataPartition(UBank_normalized$Personal,p=0.5, list=FALSE)
TrainingData = UBank_normalized[Bank_Partition,]
TestValidData = UBank_normalized[-Bank_Partition,]
Test_Index = createDataPartition(TestValidData$Personal.Loan, p=0.6, list=FALSE)
ValidationData = TestValidData[Test_Index,]
Test_Data = TestValidData[-Test_Index,]

```

```{r}
Training_Predictors<-TrainingData[,-10]
Test_Predictors<-Test_Data[,-10]
Validation_Predictors<-ValidationData[,-10]
Training_labels <-TrainingData[,10]
Test_labels <-Test_Data[,10]
Validation_labels <-ValidationData[,10]
Training_labels=as.factor(Training_labels)
Test_labels<-as.factor(Test_labels)
Validation_labels=as.factor(Validation_labels)
```
Confusion Matrix on Training data 
```{r}
knn.pred5 <- knn(Training_Predictors, Test_Predictors , cl=Training_labels, k=1, prob = TRUE)
confusionMatrix(knn.pred5,Test_labels)
```

Confusion Matrix on Validation data 
```{r}
knn.pred6 <- knn(Validation_Predictors, Test_Predictors, cl=Validation_labels, k=1, prob = TRUE)
confusionMatrix(knn.pred6,Test_labels)

```

